Foregrounds in Wide--Field Redshifted 21~cm Power Spectra

### 2015-02-23 ms.tex
Make minor edits throughout the paper
from a7b9a1f:
============

In ms.tex,

(*) Make minor edits throughout the paper

The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  \tilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}
The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units as Jy~Hz.

@@

We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach. Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$--axis. $\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.

@@

With $\hat{\boldsymbol{s}}\equiv (l,m,n)$, equation~\ref{eqn:obsvis} can be written as \citep{tay99,tho01}:
\begin{align}\label{eqn:obsvis-lmn}
  V_b(f) &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m, 
\end{align}
where, $l$, $m$, and $n$ denote the direction cosines toward east, north, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
\begin{align}\label{eqn:solid-angle}
  \dif\Omega = \frac{\dif l\,\dif m}{\sqrt{1-l^2-m^2}}.
\end{align}

@@

\begin{figure*}[htb]
\centering
% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure4a.eps}} \\
\subfloat[][Delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure4b.eps}}
\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith. Antenna shapes used are: dipole (left), phased array (middle), and dish (right). Refer to \S\ref{sec:instrument_model} for details of the antenna models. The strength and occupancy of the power patterns are correlated with those of delay power spectra. White dotted lines in the delay power spectra mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish. The phased array has intermediate levels of contamination. Foreground power close to the horizon delay limits in all three cases are significant even on long baselines. The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales. The amplitude of this feature strongly depends on the shape of the antenna element. It is highest for a dipole (which has a strong response near the horizon) and least for a dish.}
\label{fig:aperture-shapes}
\end{figure*}

@@

We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta \tilde{V} &= \sqrt{N_\textrm{ch}}\,\Delta V \Delta f, \quad\textrm{and} \\
  \Delta V &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.

@@

The most prominent signature seen in the {\it off--zenith} pointing (top left panel of Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern. It appears as a bright branch near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The bright signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.

@@

When not dominated by the bright emission from the Galactic center, the delay power spectrum of the combined foreground model is composed of diffuse and compact emission, both of which are significant. This is illustrated by a more detailed examination of the {\it zenith} pointing in our study.

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the delay power spectra plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines.

@@

Near the horizon delay limits, the diffuse component is brighter relative to the compact component. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at $\tau=0$. In fact, even on widely spaced antennas, diffuse emission in the delay power spectrum near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude. This feature was described in \S\ref{sec:diffuse-long-baselines}. We attribute this to wide--field measurement effects discussed in \S\ref{sec:wide-field}.

@@

The Galactic center in the {\it off--zenith} pointing is one such example already available in our study. Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact component, {\it bottom:} diffuse component) in this pointing. The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon. Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented northward (top panel) and eastward (bottom panel). Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom. The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay power spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}). Consequently, the spillover caused by the instrument's spectral transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe in eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.

\begin{figure*}[htb]
\centering
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figure9a.eps}}
\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figure9b.eps}}
\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figure9c.eps}}
\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}). The Galactic center is very prominent in diffuse emission on the west co--located with a bright secondary lobe of the power pattern. (b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations. Delays vary linearly with antenna spacing length. Color scale shown is for a 100~m antenna spacing. The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings. (c): Simulated delay power spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}). White lines denote horizon delay limits. The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings. These are also the most severely contaminated by foreground spillover. The northward antenna spacings, on the other hand, are the least contaminated.}
\label{fig:breakup}
\end{figure*}

@@

In other words, Figure~\ref{fig:screening} demonstrates the progress in foreground mitigation as orientation and maximum length of discarded antenna spacings are varied. The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$. It is seen that foreground contamination can be mitigated by a factor between $\sim 2$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 100$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg). The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m. Discarding antenna spacings with lengths $|\boldsymbol{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%. On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings. Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings. These results use the MWA as an example. In general, the array layout will play a significant role in determining such thresholds.

In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme. For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}. It could also be used to downweight or flag contaminated baselines in imaging applications. This technique provides a very simple and yet effective tool in adding a layer of control to mitigate effects of foreground contamination in EoR data analysis.

@@

Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA to within expected uncertainties in foreground models.

Our simulations enable us to identify numerous signatures of different components of foreground emission seen in the delay spectra. We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern, and foreground parameters such as the nature of emission, spectral index, etc.

The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing. It manifests itself near the negative horizon delay limit in the delay power spectrum on antenna spacings oriented eastward.

Diffuse emission in the primary field of view is prominent on shorter antenna spacings. However, it is also prominent near the horizon limits even on wide antenna spacings --- an effect of the wide--field nature of the measurement. On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}. Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified. In general, delay power spectrum signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in the {\it pitchfork} signature. This will be distinctly visible when the thermal noise floor is sufficiently lowered, as larger volumes of data are processed.

@@

Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper. $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at $\tau>0$ and $\tau<0$, respectively. Similarly, an object in the northern sky will appear at $\tau>0$ for an antenna spacing oriented northward. For all observations used in this study, we use zenith as the phase center, for which $\tau\equiv 0$.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{delay_cartoon.eps}
% \includegraphics[width=\linewidth]{fig1.eps}
% \includegraphics[width=\linewidth]{figure1.eps}
\includegraphics[width=\linewidth]{figureA1.eps}
\caption{Illustration of radio interferometer delays and conventions used in this paper. Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane. The local meridian, local north and local east are shown for reference. Points labeled as `z', `n', `s', `e' and `w' on the celestial sphere denote zenith, northward, southward, eastward and westward positions, respectively. $\tau_\textrm{z}$, $\tau_\textrm{n}$, $\tau_\textrm{s}$, $\tau_\textrm{e}$ and $\tau_\textrm{w}$ denote the respective delays measured between A$_1$ and A$_2$. Throughout this paper, zenith is chosen as the phase center. Hence, $\tau_\textrm{z}\equiv 0$. If $\boldsymbol{b}$ is oriented eastward as shown, then $\tau_\textrm{e} > 0$, $\tau_\textrm{w} < 0$, and $\tau_\textrm{n}=\tau_\textrm{s}=0$. Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$}. \label{fig:delay-cartoon}
\end{figure}

@@

Equation~\ref{eqn:obsvis} can be equivalently expressed as:
\begin{align}\label{eqn:vis}
  V_u(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
\end{align}
where, $\hat{\boldsymbol{s}}$ is measured with reference to a location on the sky referred to as the {\it phase center}, and $\boldsymbol{u}\equiv (u,v,w)$ denotes the spatial frequency vector. $w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it. For measurements that are constrained to lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector. Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as: 
\begin{align}\label{eqn:k-perp}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi\boldsymbol{u}}{D}, 
\end{align}
where, $D\equiv D(z)$ is the transverse comoving distance at redshift $z$.

>>>>

The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  \tilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}
The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units of Jy~Hz.

@@

We give a brief overview of some parameters of Fourier space which are generic to all experiments that use a similar approach. Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$--axis. $\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.

@@

With $\hat{\boldsymbol{s}}\equiv (l,m,n)$, equation~\ref{eqn:obsvis} can be written as \citep{tay99,tho01}:
\begin{align}\label{eqn:obsvis-lmn}
  V_b(f) &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m, 
\end{align}
where, $l$, $m$, and $n$ denote the direction cosines toward East, North, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
\begin{align}\label{eqn:solid-angle}
  \dif\Omega = \frac{\dif l\,\dif m}{\sqrt{1-l^2-m^2}}.
\end{align}

@@

\begin{figure*}[htb]
\centering
% \subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure5a.eps}} \\
% \subfloat[][delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure5b.eps}}
\subfloat[][Power patterns]{\label{fig:power-patterns}\includegraphics[width=\linewidth]{figure4a.eps}} \\
\subfloat[][Delay power spectra]{\label{fig:antennas-delay-spectra}\includegraphics[width=\linewidth]{figure4b.eps}}
\caption{Power patterns (top panels) and simulated delay power spectra (bottom panels) for different antenna shapes at 185~MHz centered on zenith. Antenna shapes used are: dipole (left), phased array (middle), and dish (right). Refer to \S\ref{sec:instrument_model} for details of the antenna models. The strength and occupancy of the power patterns are correlated with those of delay power spectra. White dotted lines in the delay power spectra mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The {\it foreground wedge} and the {\it EoR window} are most severely contaminated in the case of the dipole while it is the least for the dish. The phased array has intermediate levels of contamination. Foreground power close to the horizon delay limits in all three cases is significant even on long baselines. The foreshortening of baselines towards the horizon makes them sensitive to foreground emission on large size scales. The amplitude of this feature strongly depends on the shape of the antenna element. It is highest for a dipole (which has a strong response near the horizon) and least for a dish.}
\label{fig:aperture-shapes}
\end{figure*}

@@

We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta \tilde{V} &= \sqrt{N_\textrm{ch}}\,\Delta V \Delta f, \quad\textrm{and} \\
  \Delta V &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise in data.

@@

The most prominent signature seen in the {\it off--zenith} pointing (top left panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern. It appears as a bright branch near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The bright signature is absent in the {\it zenith} pointing (bottom left panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.

@@

When not dominated by the bright emission from the Galactic center, the delay power spectrum of the combined foreground model is composed of diffuse and compact emission both of which are significant. This is illustrated by a detailed examination of the {\it zenith} pointing.

Figure~\ref{fig:pitchfork-baselines} shows delay power spectra of three antenna pairs of different spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the delay power spectra plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines.

@@

Near the horizon delay limits, the diffuse component is brighter relative to the compact component. Here, diffuse emission does not decrease as rapidly with increasing antenna spacing as was seen at $\tau=0$. In fact, even on widely spaced antennas, diffuse emission in the delay power spectrum near the horizon delay limits exceeds that in the primary lobe by about three orders of magnitude. This feature is described in \S\ref{sec:diffuse-long-baselines} and attributed to wide--field measurement effects discussed in \S\ref{sec:wide-field}.

@@

The Galactic center in the {\it off--zenith} pointing is one such example already available in our study. Figure~\ref{fig:sky-model} shows the sky model ({\it top:} compact component, {\it bottom:} diffuse component) in this pointing. The Galactic center is the most dominant source of foreground contamination from the diffuse sky model and is co--located with a bright secondary lobe of the power pattern near the western horizon. Figure~\ref{fig:delay-map} shows the sky mapped to delays registered by the baseline vectors, of length 100~m for instance, oriented towards North (top panel) and East (bottom panel). Figure~\ref{fig:baseline-breakup} shows the delay spectra on baselines oriented northward ($67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$) at the top and eastward ($-22.5\arcdeg~\le \theta_\textrm{b} < 22.5\arcdeg$) at the bottom. The Galactic center manifests itself most distinctly near the negative horizon delay limit on short eastward baselines in the delay power spectrum (bottom panel of Figure~\ref{fig:baseline-breakup}). Consequently, the spillover caused by the instrument's spectral transfer function from the {\it foreground wedge} into the {\it EoR window} affects the northward baselines the least and is most severe on eastward baselines (particularly the short ones) evident by the bright vertical stripes of foreground contamination.

\begin{figure*}[htb]
\centering
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figure9a.eps}}
\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figure9b.eps}}
\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figure9c.eps}}
\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}). The Galactic center is very prominent in diffuse emission on the West co--located with a bright secondary lobe of the power pattern. (b): Sky hemisphere mapped to delays observed on antenna spacings oriented North ({\it top}) and East ({\it bottom}). Delays vary linearly with antenna spacing length. Color scale shown is for a 100~m antenna spacing. The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings. (c): Simulated delay power spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}). White lines denote horizon delay limits. The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings. These are also the most severely contaminated by foreground spillover. The northward antenna spacings, on the other hand, are the least contaminated.}
\label{fig:breakup}
\end{figure*}

@@

In other words, Figure~\ref{fig:screening} demonstrates the progress in foreground mitigation as orientation and maximum length of discarded antenna spacings are varied. The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$. It is seen that foreground contamination can be mitigated by a factor between $\sim 2$ ($|\theta_\textrm{b}|\le$~7.5\arcdeg) and $\sim 100$ ($|\theta_\textrm{b}|\le$~22.5\arcdeg). The latter limit is achieved with a mere 5\% loss of data for $|\boldsymbol{b}|_\textrm{max}\simeq30$~m. Discarding antenna spacings with lengths $|\boldsymbol{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%.

If there was a bright point source in place of the Galactic center, it will give rise to foreground contamination even on longer antenna spacings. Such cases will necessitate discarding more or all of the eastward antenna spacings. The MWA was used as an example. % These results use the MWA as an example. 
In general, the direction, strength and type of emission, and the array layout will determine such thresholds.

In principle, instead of discarding selected antenna spacings altogether, we could down--weight them based on an optimal scheme. For instance, the estimates of covariance computed from the delay transform bins can be naturally fed into the covariance--weighted power spectrum estimation techniques \citep{liu14a,liu14b}. It could also be used to downweight or flag contaminated baselines in imaging applications. This technique provides a very simple and yet effective tool to suppress the effects of foreground contamination in EoR data analysis.

@@

Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission and bright point sources, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA to within expected uncertainties in foreground models.

Our simulations enable us to identify numerous signatures of different components of foreground emission seen in the delay spectra. We establish the relationship between these signatures and observing parameters such as antenna pointing and LST, instrument parameters such as antenna power pattern, and foreground parameters such as type of emission, etc.

The bright Galactic center at the edge of the western horizon co--located with one of the far secondary lobes of MWA tile power pattern is the brightest source of foreground contamination in the {\it off--zenith} pointing. It manifests itself near the negative horizon delay limit in the delay power spectrum on eastward antenna spacings.

Diffuse emission in the primary field of view is prominent on shorter antenna spacings. However, it is also prominent near the horizon limits even on wide antenna spacings --- an effect of the wide--field nature of the measurement. On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}. Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified. In general, delay power spectrum signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in a characteristic {\it pitchfork} signature. This will be evident when thermal noise is sufficiently lowered, as larger volumes of data are processed.

@@

Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper. $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local East, North (along local meridian) and upward (zenith) directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at $\tau>0$ and $\tau<0$, respectively. Similarly, an object in the northern sky will appear at $\tau>0$ for an antenna spacing oriented northward. For all observations used in this study, we use zenith as the phase center, for which $\tau\equiv 0$.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{delay_cartoon.eps}
% \includegraphics[width=\linewidth]{fig1.eps}
% \includegraphics[width=\linewidth]{figure1.eps}
\includegraphics[width=\linewidth]{figureA1.eps}
\caption{Radio interferometer delay conventions used in this paper. Two antennas labeled as A$_1$ and A$_2$ are separated by vector $\boldsymbol{b}$ on the local tangent plane. The local meridian, local North and local East are shown for reference. Points labeled as `z', `n', `s', `e' and `w' on the celestial sphere denote zenith, northward, southward, eastward and westward positions, respectively. $\tau_\textrm{z}$, $\tau_\textrm{n}$, $\tau_\textrm{s}$, $\tau_\textrm{e}$ and $\tau_\textrm{w}$ denote the respective delays measured between A$_1$ and A$_2$. Throughout this paper, zenith is chosen as the phase center. Hence, $\tau_\textrm{z}\equiv 0$. If $\boldsymbol{b}$ is oriented eastward as shown, then $\tau_\textrm{e} > 0$, $\tau_\textrm{w} < 0$, and $\tau_\textrm{n}=\tau_\textrm{s}=0$. Conversely, if $\boldsymbol{b}$ is oriented northward (not shown here), then $\tau_\textrm{n} > 0$, $\tau_\textrm{s} < 0$, and $\tau_\textrm{e}=\tau_\textrm{w}=0$.} \label{fig:delay-cartoon}
\end{figure}

@@

Equation~\ref{eqn:obsvis} can be equivalently expressed as:
\begin{align}\label{eqn:vis}
  V_u(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
\end{align}
where, $\hat{\boldsymbol{s}}$ is measured with reference to a location on the sky referred to as the {\it phase center}, and $\boldsymbol{u}\equiv (u,v,w)$ denotes the spatial frequency vector. $w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it. For measurements that lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector. Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as: 
\begin{align}\label{eqn:k-perp}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi\boldsymbol{u}}{D}, 
\end{align}
where, $D\equiv D(z)$ is the transverse comoving distance at redshift $z$.
